---
title: "A3: SPL Library Checkouts"
output: html_document
---

## Author
Amy Ho

### Introduction

> The dataset that I chose focused on items checked out at least ten times a month from 2017 to 2023 by the Seattle Public Library. By analyzing numerous variables, this report explores where trends of titles, subjects, and material types can fall or increase. Collection of checkouts by item type have revealed the rise of Ebook checkouts. Hence, comparing two subjects of fiction and nonfiction display the surge of checkouts and demand for a subject. We can learn about the popularity of titles by the same author throughout the years. Data on titles by certain authors can teach us about reading preferences of various demographics. 

</br>

### Summary Information
```{r, echo=FALSE, message=FALSE, code = readLines("summary.R")}
source("summary.R")
```

> According to the data findings, total fiction checkouts had a total of `r total_fiction` (5282) checkouts, whereas total nonfiction checkouts had a total of `r total_nonfiction` (2548) checkouts. Likewise, the year 2018 had the most total book checkouts of `r most_book_checkout` (1,527,112), but the year 2023 had the least total book checkouts of `r least_book_checkout` (85,080), a significant decrease in book checkouts. Overall, book checkouts were `r sum_book` (6,397,967) and Ebook checkouts were `r sum_ebook` (5,053,218), tying them in close for popularity of material type. Book checkouts remain the most checked out material type despite a decrease checkouts every year. 

</br>

### The Dataset

Who collected/published the data?  

> This dataset is from the Seattle Public Library that includes the monthly counts and checkouts by title for different material types. The chosen dataset in this report is for items checked out at least 10 tens a month from 2017 to 2023. 

What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?

>  The parameters of the data includes twelve columns from usage class between physical or digital, checkout type of horizon or overdrive, material type of books, ebooks, audiobooks, and many more. The range of data is from 2017 to 2023 of each material type. It also includes author names, publishers, and title names. 

How was the data collected or generated?  

> The data comes from metadata of current and historical sources dating back to 2005 for the main datatset. Digital items come from media vendors such as Overdrive, but data past 2016 is provided by Horizon. Data is refreshed monthly and only counts the initial checkout.

Why was the data collected?

> This data is collected and used as an open data platform to keep track of checkouts by title, but also serves as storing metadata. It is important to collect this data for public access for finding out what to read, transparency, understanding trends, and allows anyone to download it free of charge. 

What, if any, ethical questions do you need to consider when working with this data?  

> Some ethical questions that can be raised while working with this set is how this data is being used or possibly sold and how there could possibly be a security issue.

What are possible limitations or problems with this data?   (at least 200 words)

> Many limitations and problems arise from the massive number of rows and data points in this collection. The "Creator" column is blank for a large number of the titles. Since digital content is compiled from a wide variety of media sources, without a Horizon record, there is no way to know what each piece is about since they were not recorded. The year publishing (e.g. 2017), year with copyright symbol, printing year, phonograph copyright symbol, publication and copyright date, intervening years, and approximate date among the many possible year formats that can be found in the publication year column. It's also possible that there are gaps, inconsistencies, and security concerns with the data.

</br>

### First Chart Representing Monthly Checkouts by J.R.R. Tolkien of Various Books
```{r chart 1, echo=FALSE, message=FALSE, warning=FALSE,code= readLines("chart1_tolkien.R")}
source("chart1_tolkien.R")
```

> In this chart, we are looking at number of monthly checkouts of J.R.R. Tolkien books. This is a line plot where the authors most popular titles stay in the same number of checkouts range. We can see in our chart that "The Hobbit" and "The Lord of the Rings, the Fellowship of the Ring" stay in a similar range. Yet, there is a sudden spike in checkouts after the year 2021 and remains in a high trend. The purpose of this chart allows us to understand the trend of books by author J.R.R. Tolkien and when popularity emerges due to a spike. Additionally, this chart identifies when "The Hobbit" becomes popular and visually seeing an odd occurence of high checkouts. A line chart shows an upward trend of "The Hobbit" in a certain time period whereas the other trend lines generally remain in the same level. These trends are in the same time period and put into a line plot as the number of checkouts are not large and there are smaller changes.

</br>

### Second Trends Over Time Chart of Fiction and Nonfiction
```{r, echo=FALSE, message=FALSE, warning=FALSE,code= readLines("chart2_genres.R")}
source("chart2_genres.R")
```

> The chart is a representation of fiction versus nonfiction checkouts over the years. While fiction checkouts start at a high quantity of checkouts, there is about a 1000 checkout difference in 2018 and 2020, but nonfiction checkouts dramatically increase in 2021 leaving more than a 2000 checkout gap and suddenly decreases. The two subjects are at an equal amount in late 2021 of roughly 800 checkouts. This chart allows us to visually compare fiction and nonfiction checkouts despite the numbers given. Without this chart, you would not be able to see that fiction checkouts dropped and nonfiction checkouts increased. We are able to visualize the trends of fiction checkouts falling whereas nonfiction trends have a sudden spike and drop. The trend shows that as fiction decreases, nonfiction increases. A line chart is chosen because it is able to display small clear changes of trends over time between two genres to make predictions about genre popularity.

</br>

### Bar Graph of Material Types from 2017 to 2023
```{r, echo=FALSE, message=FALSE, warning=FALSE, code= readLines("chart3_bar.R")}
source("chart3_bar.R")
```

> This horizontal bar graph counts the total of checkouts by material item type. The number of book checkouts decrease every year, but ebooks increase every year showing how ebooks have gained popularity and are nearly the same amount as book checkouts.This data can be analyzed to see where item trends start to increase in comparison to other material types. The purpose of this graph allows us to see separated bars of material type of checkouts from 2017 from bottom to top. Hence, if we follow the trend from the bottom of the graph to the top, we can see the direction of material types changing over time and helps us predict when an item type is developing. We can see that ebooks are gradually catching up to book checkouts. A bar graph makes it easy to compare between material type in the dataset, where the y-axis represents material types and the x-axis represents total number of checkouts over time. A horizontal bar graph allows room to see labels on the y-axis and is effective for larger numbers and changes. Thus, we are able to compare several categories and be able to identify trends of popularity of material type. 

